Library Learning in Neurally-Guided Bayesian Program Learning

### 2018-09-22 nips_2018.tex
ec2

\newcommand{\system}{\textsc{SCC} }
\newcommand{\systemEnding}{\textsc{SCC}}

@@

\author{
  Kevin Ellis\\MIT\And Lucas Morales\\MIT\And Mathias Sabl\'e Meyer\\France
  \AND
  Armando Solar-Lezama\\MIT\And Joshua B. Tenenbaum\\MIT
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

Our \system (\textbf{S}earch/\textbf{C}ompress/\textbf{C}ompile)
algorithm incorporates these insights by iterating through three steps. 
The \textbf{Search} step takes a given set of \textbf{tasks}, typically
several hundred, and searches for compact programs that solve these tasks
guided by the current DSL and neural network.
The \textbf{Compress} step grows the library (or DSL) of
domain-specific subroutines which allow the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during search, compressing
them to distill out common code fragments across successful programs.
The \textbf{Compile} step improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.%%  (CITE Goodman and Stuhlmuller,
%% and others too on amortized inference.)

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t.\ the frontiers (Search) and the DSL (Compression):
\\\noindent \textbf{Program Search: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\subsection{Searching for Programs}\label{explorationSection}

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{79\%}\footnotemark & 
  76\% &71\%&35\%&62\%&37\%\\
  Solve time&  4.1s&5.8s&10.6s&34.7s&43.4s&20.2s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% held-out test tasks solved. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first search step.
&
%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.

>>>>

\newcommand{\system}{\textsc{EC$^2$} }
\newcommand{\systemEnding}{\textsc{EC$^2$}}

@@

\author{
  Kevin Ellis\\MIT\\\texttt{ellisk@mit.edu}\And Lucas Morales\\MIT\And Mathias Sabl\'e-Meyer\\ENS Paris-Saclay
  \AND
  Armando Solar-Lezama\\MIT\And Joshua B. Tenenbaum\\MIT
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \AND
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
  %% \And
  %% Coauthor \\
  %% Affiliation \\
  %% Address \\
  %% \texttt{email} \\
}

@@

Our \system (ECC, for \textbf{E}xplore/\textbf{C}ompress/\textbf{C}ompile)
algorithm incorporates these insights by iterating through three steps. 
The \textbf{Explore} step takes a given set of \textbf{tasks}, typically
several hundred, and explores the space of programs, searching for compact programs that solve these tasks,
guided by the current DSL and neural network.
The \textbf{Compress} step grows the library (or DSL) of
domain-specific subroutines which allow the agent to more compactly
write programs in the domain; it modifies the structure of the DSL by
discovering regularities across programs found during exploration, compressing
them to distill out common code fragments across successful programs.
The \textbf{Compile} step improves the search procedure by training a neural network to
write programs in the current DSL, in the spirit of ``amortized'' or
``compiled'' inference~\cite{le2016inference}.%%  (CITE Goodman and Stuhlmuller,
%% and others too on amortized inference.)

@@

%% In general this marginalization over $\theta$ is intractable, so we make an AIC-style approximation\footnote{Sec.~\ref{grammarInductionSection} explains that $\mathcal{D}$ is a context-sensitive grammar.
%% Conventional natural-language processing (NLP) approaches to using variational inference to lower bound the marginal over $\theta$ do not apply in our setting.}, $A\approx \log\probability[\mathcal{D}|X] $:
%% \begin{align}
%%   A =   \log \probability[\mathcal{D}] + \argmax_{\theta}& \sum_{x\in X}\log \sum_p\probability[x|p]\probability[p|\mathcal{D},\theta]\nonumber\\
%% &+  \log P(\theta|\mathcal{D}) - ||\theta||_0 \label{AIC}
%%   \end{align}
\system does approximate MAP inference  by maximizing this lower bound on the joint probability,
alternating maximization w.r.t.\ the frontiers (Explore) and the DSL (Compression):
\\\noindent \textbf{Explore: Maxing $\lowerBound$ w.r.t.\ the frontiers.} Here $(\mathcal{D},\theta)$ is fixed and we
want to find new programs to add to  the frontiers so that $\lowerBound$ increases the most.
$\lowerBound$ most increases by finding programs where $\probability[x|p]\probability[p|\mathcal{D},\theta]$
is large.
%% which we can accomplish by adding new programs to the frontiers means searching for new programs $p$ for task $x$
%% where  is large.
\\\noindent \textbf{DSL Induction: Maxing $\int \lowerBound\;\mathrm{d}\theta$ w.r.t.\ the DSL.} Here $\left\{\mathcal{F}_x \right\}_{x\in X}$ is held fixed, and so we can evaluate $\lowerBound$. Now the problem is that of searching the discrete space of DSLs and finding one maximizing $\int \lowerBound\;\mathrm{d}\theta$.
Once we have a DSL $\mathcal{D}$ we can update $\theta$ to $\argmax_\theta \lowerBound(\mathcal{D},\theta,\left\{\mathcal{F}_x \right\})$.

@@

\subsection{Exploration: Searching for Programs}\label{explorationSection}

@@

We compare with ablations of our model on held out tasks.
The purpose of this ablation study is 
both to examine the role of each component of \systemEnding,
as well as to compare with
prior approaches in the literature:
a head-to-head
comparison of program synthesizers is complicated by the fact that
each system, including ours, makes idiosyncratic 
assumptions about the space of programs and the statement of tasks.
\begin{wraptable}{r}{0.55\textwidth}
\vspace{-5pt}
%% This radically shrinks the table
\tabcolsep=2.5pt
\renewcommand{\arraystretch}{0.5}
\begin{tabular}{lcccccc}
  \toprule& Ours& 
    No NN & SE&NPS & PCFG & Enum \\
  \midrule
  \multicolumn{7}{c}{\emph{List Processing}}\\
  \midrule
  \% solved&\textbf{79\%}\footnotemark & 
  94\% &71\%&35\%&62\%&37\%\\
  Solve time&  88s&6s&11s&35s&44s&20s\\
  \midrule
  \multicolumn{7}{c}{\emph{Text Editing}}\\
  \midrule
  \% solved&\textbf{74\%} &43\% &30\%&33\%&0\%&4\%\\
  Solve time& 29s&65s&38s&80s&--&235s\\
  \midrule
  \multicolumn{7}{c}{\emph{Symbolic Regression}}\\
  \midrule
  \% solved&   \textbf{84\% }&75\%&62\%&38\%&38\%&37\% \\
  Solve time&  24s& 40s  &28s&31s&55s&29s\\
  \bottomrule
  \end{tabular}
\caption{\% held-out test tasks solved. Solve time: averaged over solved
  tasks.  %SE: NPS: trained like RobustFill/DeepCoder.  PCFG: model w/o
  %structure learning.  Enum: model w/o any learning.  %% MDL:
  %% $-\expect\left[\probability[x|p] \right]$. For domains other than
  %% symbolic regression MDL is 0 nats.
  %% For symbolic regression MDL is a
  %% proxy for \# of continuous parameters.
}\vspace{-0.5cm}\label{baselineComparisons} \end{wraptable}
Nevertheless, much prior work can be modeled within our setup. 
We compare with the following ablations (Tbl~\ref{baselineComparisons};
Fig~\ref{learningCurves}):
\\\noindent \textbf{No NN:} lesions the recognition model.
\\\noindent \textbf{NPS}, which does not learn the DSL,
instead learning the recognition model
from samples drawn from the fixed DSL.
We call this NPS (Neural Program Synthesis)
because this is closest to how
RobustFill~\cite{devlin2017robustfill} and DeepCoder~\cite{balog2016deepcoder} are trained.
\\\noindent \textbf{SE}, which lesions the recognition model and restricts the DSL  learning algorithm to
only add \textbf{S}ub\textbf{E}xpressions of programs in the frontiers to the DSL. This is how most prior approaches have learned libraries of functions~\cite{Dechter:2013:BLV:2540128.2540316,DBLP:conf/icml/LiangJK10,DBLP:conf/ecai/LinDETM14}.
\\\noindent \textbf{PCFG}, which lesions the recognition model and does not learn the DSL,
but instead learns the parameters of the DSL ($\theta$), learning the parameters of a PCFG while not learning any of the structure.
\\\noindent \textbf{Enum}, which enumerates a frontier without any learning --- equivalently, our first exploration step.
&
%For each domain,
%% We are interested both in how many tasks the
%% agent can solve and how quickly it can find those solutions.
%% Tbl.~\ref{baselineComparisons}
%% compares our model against these alternatives.
%% We consistently
%% improve on the baselines,
%% and find that lesioning the recognition model
%% and lesioning it also slows down the convergence of the algorithm,
%% taking more iterations to reach a given number of tasks solved (Fig.~\ref{learningCurves}).
%% This supports a view of the recognition model as a way of amortizing the cost of search.
